The Analog Method for Precipitation Downscaling: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-07 Moving temporal window - HESS/template.tex
Minor edits.
Working on the HESS paper corrections.

%% Journal Abbreviations (Please use the same for Discussion Papers and Final Revised Papers)

% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Advances in Geosciences (adgeo)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% ASTRA Proceedings (ap)
% Atmospheric Measurement Techniques (amt)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Biogeosciences (bg)
% Climate of the Past (cp)
% Drinking Water Engineering and Science (dwes)
% Earth System Dynamics (esd)
% Earth Surface Dynamics (esurf)
% Earth System Science Data (essd)
% Fossil Record (fr)
% Geographica Helvetica (gh)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% Geothermal Energy Science (gtes)
% Hydrology and Earth System Sciences (hess)
% History of Geo- and Space Sciences (hgss)
% Journal of Sensors and Sensor Systems (jsss)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Primate Biology (pb)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Web Ecology (we)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

@@

% \Author[affil]{given_name}{surname}

@@

\affil[1]{University of Lausanne, Lausanne, Switzerland}
\affil[2]{Institute of Geography \& Oeschger Centre for Climate Change Research, University of Bern, Bern, Switzerland}
\affil[3]{Universit\'{e} de Grenoble-Alpes, LTHE, Grenoble, France}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

@@

@@

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the Continuous Ranked Probability Score \citep[$S_{\text{CRP}}$,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. It is defined as follows:

@@

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the prediction skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (Sect. \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (Sect. \ref{sec:method}).

The MTW is explained in Sect. \ref{sec:method}. As this method requires sub-daily time series, the available archive length is reduced compared to more common daily measurements, which has some consequences on the performance score (Sect. \ref{sec:archive_reduction}). The benefit of introducing a MTW was assessed first in regards to the analogy criteria improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (Sect. \ref{sec:attempts}).

\section{Data and method}

Predictors are extracted from the NCEP/NCAR reanalysis I \citep{Kalnay1996} dataset with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree.

\subsection{The moving time window (MTW) approach}
\label{sec:method}

@@

Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated using AtmoSwing \citep{Horton2016} on the original archive (Tables \ref{table:performance}) and will be used thereafter.

@@

\section{Results and discussion}

@@

@@

\subsubsection{Influence of the weather situation}
\label{sec:influence_precip}

@@

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of $S_{\text{TW}}$ for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

\subsubsection{Seasonal effect}
\label{sec:seasonal_effect}

A seasonal effect can be observed on the reduction of the $S_{\text{TW}}$ criteria due to the MTW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogs are essentially found for the same hours or time windows.

@@

A systematic improvement of the $S_{\text{TW}}$ values was previously observed. However, finding better analog situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analog dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC -- 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

\subsubsection{Improvement by precipitation classes}
\label{sec:improvement_CRPSS_precip_threshold}

\subsubsection{Effect of recalibrating the parameters}
\label{sec:recalibration}

After recalibrating, some changes in the optimal number of analogs can be observed for both 2Z and 2Z-2MI methods (Table \ref{table:nb_analogs_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogs increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogs. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the $S_{\text{TW}}$ criteria grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

The values of the $S_{\text{SCRP}}$ scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated.

\section{Application attempts on the full archive}
\label{sec:attempts}

The improvement provided by the MTW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving 24~h-totals, different disaggregation approaches of the daily time series were assessed and are presented in the following sections.

\subsection{Proportional distribution}

@@

\subsection{Use of a proxy variable}

@@

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5--7.5° E and 45--47.5° N), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

The results of the correlation analyses (on non-zero values) between the new series reconstituted using the proxy and the actual series are illustrated in Table \ref{table:proxy_correlations} for Zermatt (on the short period, 1982--2007). The best proxy is the precipitable water at 45° N and 7.5° E, with a time offset of -6~h. The determination coefficient of 0.721 is higher than that of the series from the proportional distribution (0.698), which confirms that a little information was added to the series of precipitation.

Table \ref{table:proxy_CRPSS} presents the $S_{\text{SCRP}}$ scores obtained by the reconstituted series using the proxy of precipitable water at the optimal point (Table \ref{table:proxy_correlations}). A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the MTW is again lost.

@@

\bibliographystyle{copernicus}
\bibliography{../_refs/_articles-moving-window}

@@

\begin{table}[htb]
	\caption{Value of the coefficient of determination between the reconstructed 6-hourly precipitation time series using the listed variables, and the accurate time series on the period 1982-2007. The grid points are the following: 1) 5° E, 47.5° N, 2) 5° E, 45° N, 3) 7.5° E, 47.5° N, 4) 7.5° E, 45° N. The highest coefficient of determination is indicated in bold.}
	\begin{center}
		\begin{tabular}{l c c c c c c}
			\hline
			\multirow{2}{*}{Variable} & \multirow{2}{*}{Point} &  \multicolumn{5}{c}{Time lapse} \\
			&  & -12 h & -6 h & 0 h & +6 h & +12 h \\ 
			\hline
			\multirow{ 4}{*}{RH1000} & 1 & 0.668 & 0.669 & 0.684 & 0.683 & 0.670 \\
			& 2 & 0.669 & 0.669 & 0.683 & 0.681 & 0.669 \\
			& 3 & 0.662 & 0.673 & 0.691 & 0.682 & 0.673 \\
			& 4 & 0.666 & 0.671 & 0.688 & 0.681 & 0.668 \\ \hline
			\multirow{ 4}{*}{RH925} & 1 & 0.672 & 0.673 & 0.684 & 0.684 & 0.675 \\
			& 2 & 0.674 & 0.674 & 0.683 & 0.682 & 0.672 \\
			& 3 & 0.662 & 0.673 & 0.691 & 0.682 & 0.673 \\
			& 4 & 0.666 & 0.671 & 0.689 & 0.681 & 0.668 \\ \hline
			\multirow{ 4}{*}{RH850} & 1 & 0.675 & 0.675 & 0.679 & 0.678 & 0.671 \\
			& 2 & 0.681 & 0.690 & 0.691 & 0.677 & 0.664 \\
			& 3 & 0.665 & 0.680 & 0.693 & 0.683 & 0.675 \\
			& 4 & 0.675 & 0.694 & 0.706 & 0.681 & 0.659 \\ \hline
			\multirow{ 4}{*}{TCW} & 1 & 0.688 & 0.687 & 0.667 & 0.655 & 0.652 \\
			& 2 & 0.697 & 0.699 & 0.669 & 0.644 & 0.644 \\
			& 3 & 0.686 & 0.708 & 0.689 & 0.655 & 0.648 \\
			& 4 & 0.696 & \textbf{0.721} & 0.696 & 0.643 & 0.636 \\ \hline
		\end{tabular}
	\end{center}
	\label{table:proxy_correlations}
\end{table}

>>>>

%% Journal Abbreviations (Please use the same for Discussion Papers and Final Revised Papers)
% Hydrology and Earth System Sciences (hess)

@@

@@

\affil[1]{Institute of Earth Sciences, University of Lausanne, Lausanne, Switzerland}
\affil[2]{Oeschger Centre for Climate Change Research \& Institute of Geography, University of Bern, Bern, Switzerland}
\affil[3]{Universit\'{e} de Grenoble-Alpes, LTHE, Grenoble, France}

@@

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the prediction skills \citep{Finet2008}.

The MTW is explained in Sect. \ref{sec:mtw}. As this method requires sub-daily time series, the available archive length is reduced compared to more common daily measurements, which has some consequences on the performance score (Sect. \ref{sec:archive_reduction}). The benefit of introducing a MTW was assessed first in regards to the analogy criteria improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion}.

%TODO: complete !!

\section{Data and methods}

Predictors are extracted from the NCEP/NCAR reanalysis I \citep{Kalnay1996} dataset with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but still often used (see discussion).

\subsection{The considered analog method}
\label{sec:analog_method}

@@

\subsection{Calibration of the analog method}
\label{sec:calibration}

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}.

\subsection{The moving time window (MTW) approach}
\label{sec:mtw}

In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (Sect. \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total.

\subsection{Performance score}
\label{sec:performance}

The score that is most often considered to assess an AM performance is the Continuous Ranked Probability Score \citep[$S_{\text{CRP}}$,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. It is defined as follows:

@@

Finding better analog situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analog dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC -- 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

\section{Results}

@@

Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (Tables \ref{table:performance}) and will be used thereafter.

@@

@@

\subsubsection{Changes per precipitation classes}
\label{sec:influence_precip}

The impact of the MTW on the analogy criteria has been decomposed per precipitation classes. The results of this analysis are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of $S_{\text{TW}}$ for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations.

\subsubsection{Changes per season}
\label{sec:seasonal_effect}

A seasonal effect can be observed on the reduction of the $S_{\text{TW}}$ criteria due to the MTW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer.

A systematic improvement of the $S_{\text{TW}}$ values was previously observed.

\subsubsection{Improvement per precipitation classes}
\label{sec:improvement_CRPSS_precip_threshold}

\subsubsection{Recalibration of the AM parameters}
\label{sec:recalibration}

After recalibrating, some changes in the optimal number of analogs can be observed for both 2Z and 2Z-2MI methods (Table \ref{table:nb_analogs_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease (not shown). More significantly, the optimal number of analogs increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

The values of the $S_{\text{SCRP}}$ scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration.

\section{Discussion}
\label{sec:discussion}

\subsection{Improvement of the atmospheric circulation analogy}

@@

\subsection{Improvement of the moisture analogy}

With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated.

\subsection{Impact of the MTW per precipitation classes}

@@

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is.

\subsection{Seasonal effect}

One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogs are essentially found for the same hours or time windows.

@@

\subsection{Increase in number of analog situations}

It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogs. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the $S_{\text{TW}}$ criteria grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

\subsection{Why not just making a 6-hourly prediction ?}

\subsection{On the use of an old reanalysis dataset}

The considered reanalysis dataset, which is the NCEP/NCAR reanalysis I \citep{Kalnay1996}, is relatively old and has a coarse resolution. Newer reanalysis dataset could have been used to achieve this study, and might have resulted in higher performance scores. However, one can assume that the gain observed here by introducing a MTW should also be found for a better reanalysis dataset, especially if the information of the atmospheric circulation is more accurate.

\subsection{Is a simple reconstructed archive useful for the MTW ?}

The improvement provided by the MTW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving 24~h-totals, different disaggregation approaches of the daily time series were assessed.

@@

\subsection{Can a proxy variable help reconstruct a better archive ?}

@@

The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point (details not shown). The best proxy for the Zermatt station is the precipitable water at 45° N and 7.5° E, with a time offset of -6~h.

Table \ref{table:proxy_CRPSS} presents the $S_{\text{SCRP}}$ scores obtained by the reconstituted series using the proxy of precipitable water. A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the MTW is again lost. A more recent reanalysis archive with more accurate moisture variables might produce better proxies.

@@

\bibliographystyle{copernicus}
\bibliography{../_refs/4_articles-2016_HESS_moving-window}

@@
